Do motif roles provide unique information about species risk of extinction?\\ \medskip Supplementary Information

### 2021-09-29 manuscript/supplement_motifs.tex
still syncing SI and main

        As with network profiles, species' motif participation may covary with other measures of network structure and species' roles within the network.
        To establish this context for motif participation, we first fit a PERMANOVA~\citep{Anderson2001} to test whether a species' overall motif participation role is related to network size, connectance, or the interaction between them.
        To avoid inflating similarity between species which do not appear in the same motifs (i.e., which have ``shared zeros''), we use Bray-Curtis dissimilarity between motif participation roles.
        Significance was calculated using 9999 permutations.
        The PERMANOVA was fit using the R~\citep{R} function `adonis' from the package \emph{vegan}~\citep{vegan}.

@@

        To test whether networks with similar motif profiles tend to have similar mean persistences among consumers, we fit a PERMANOVA relating Bray-Curtis dissimilarity in motif profiles to the mean persistence of all consumers in a network (averaged across all levels of disturbance to basal resources).
        We then tested whether dispersion was homogeneous across levels of persistence using an ANOVA and whether there was a linear relationship between persistence (rounded to three decimal places) and variability of motif profiles using a linear regression.
        In case high or low mean persistences are associated with more variable network structure, we calculated dispersion of motif structure for each level of mean persistence (rounded to three decimal places to allow multiple networks to have the same value of mean consumer persistence and treated as a categorical variable). 
        All statistical tests were conducted as when relating motif profiles to global network structure.

        To supplement these overall tests, we also fit four regressions testing whether persistence had a linear relationship to the proportion of a network profile made up of each motif, and whether this varied with the level of disturbance.
        Each regression included the proportion of a motif, disturbance, and their interaction as predictors.
        All regressions were fit using the R~\citep{R} base function `lm'.

@@

        As well as affecting persistence directly, global network structure (i.e., size and connectance) may affect the meso-scale structure of the network.
        This is especially likely since we include only networks which can retain all initial species after 1000 rounds of simulated Lotka-Volterra population dynamics~\citep{Cirtwill2021_inprep}:  some meso-scale structures might be more likely to allow all species to persist than others in networks with different size and connectance constraints.
        We test whether networks with the the same size and connectance have more similar motif profiles using a PERMANOVA~\citep{Anderson2001} of Bray-Curtis dissimilarity among motif profiles against network size, connectance, and their interaction.
        Significance was calculated based on 9999 permutations.
        PERMANOVA tests may give false positive results if variability is not homogeneous among groups (i.e., levels of network size or connectance).
        To test whether this applies in our case, we fit an ANOVA of the dispersion of motif profiles about the centroid for each combination of network size and connectance. 
        We fit the PERMANOVA using the function `adonis', calculated dispersion using the function `betadisper', and fit the ANOVA using the function `anova', all from the R~\citep{R} package \emph{vegan}~\citep{vegan}.
        To gain a more detailed picture of how the proportion of each motif may vary with global network structure, we also fit four regressions (one per motif) relating the proportion of a motif to network size, connectance, and their interaction.
        We fit each regression using the R~\citep{R} base function `lm'.

@@

>>>>

        As with network profiles, species' motif participation may covary with other measures of network structure and species' roles within the network.
        To establish this context for motif participation, we first fit a PERMANOVA~\citep{Anderson2001} to test whether a species' overall motif participation role is related to network size, connectance, or the interaction between them.
        To avoid inflating similarity between species which do not appear in the same motifs (i.e., which have ``shared zeros''), we use Bray-Curtis dissimilarity between motif participation roles.
        Significance was calculated using 9999 permutations.

        PERMANOVA tests may give false positive results if variability is not homogeneous among groups (i.e., levels of network size or connectance).
        To test whether this applies in our case, we fit an ANOVA of the dispersion of motif participation roles about the centroid for each combination of network size and connectance. 
        We then regressed dispersion against persistence to test whether networks with higher or lower persistence had more consistent motif profiles.
        We fit the PERMANOVA using the function `adonis', calculated dispersion using the function `betadisper', and fit the ANOVA using the function `anova', all from the R~\citep{R} package \emph{vegan}~\citep{vegan}.
        We fit the regression using the R~\citep{R} base function `lm'.

@@

        To test whether networks with similar motif profiles tend to have similar mean persistences among consumers, we fit a PERMANOVA relating Bray-Curtis dissimilarity in motif profiles to the mean persistence of all consumers in a network (averaged across all levels of disturbance to basal resources).
        As with the PERMANOVA relating motif participation to network properties, we fit the PERMANOVA using the R~\citep{R} function `vegan' from the package \emph{adonis}~\citep{adonis} and calculated significance based on 9999 permutations.

        We then tested whether dispersion was homogeneous across levels of persistence using an ANOVA and whether there was a linear relationship between persistence (rounded to three decimal places) and variability of motif profiles using a linear regression.
        Dispersion was calculated using the R~\citep{R} function `betadisper' from the package \emph{vegan}~\citep{vegan}.
        In case high or low mean persistences are associated with more variable network structure, we calculated dispersion of motif structure for each level of mean persistence (rounded to three decimal places to allow multiple networks to have the same value of mean consumer persistence and treated as a categorical variable).

        To supplement these overall tests, we also fit four linear models relating mean persistence to network motif profiles, disturbance, and the interaction between them:

            \begin{equation}
                \Psi_{ik} \approx \Bar{\rho}_{i} + \pi_{disturbed_k} + \Bar{\rho}_{i}\pi_{disturbed_k} ,
                \label{netpropeq}
            \end{equation}

        where $\Bar{\rho}_{i}$ is the proportion of the focal motif in the network's motif profile and all other symbols are as in equation~\ref{propreq}. 
        All regressions were fit using the R~\citep{R} base function `lm'.

@@

        As well as affecting persistence directly, global network structure (i.e., size and connectance) may affect the meso-scale structure of the network.
        This is especially likely since we include only networks which can retain all initial species after 1000 rounds of simulated Lotka-Volterra population dynamics~\citep{Cirtwill2021_inprep}:  some meso-scale structures might be more likely to allow all species to persist than others in networks with different size and connectance constraints.
        We test whether networks with the the same size and connectance have more similar motif profiles using a PERMANOVA~\citep{Anderson2001} of Bray-Curtis dissimilarity among motif profiles against network size, connectance, and their interaction.
        Significance was calculated based on 9999 permutations.
        PERMANOVA tests may give false positive results if variability is not homogeneous among groups (i.e., levels of network size or connectance).
        To test whether this applies in our case, we fit an ANOVA of the dispersion of motif profiles about the centroid for each combination of network size and connectance. 
        We fit the PERMANOVA using the function `adonis', calculated dispersion using the function `betadisper', and fit the ANOVA using the function `anova', all from the R~\citep{R} package \emph{vegan}~\citep{vegan}.

        To gain a more detailed picture of how the proportion of each motif may vary with global network structure, we also fit four regressions (one per motif) relating the proportion of a motif to network size, connectance, and their interaction:

        \begin{equation}
            \Bar{\rho_{i}} \approx \Sigma_{i} + \zeta_{i} + \Sigma_{i}\zeta_{i}
        \end{equation}
        where all symbols are as in equations~\ref{propreq} and~\ref{netpropeq}.
        We fit each regression using the R~\citep{R} base function `lm'.

@@

        S, C for random effects. $\Sigma$ and $\zeta$ for fixed effects.
